---
title: "WUP_04"
author: "Jayline Zelaya"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSS=read.csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The response variable is WORDUM which is numerical. The predictor variable is DEGREE which is categorical. Therefore we use numerical~cat or ChatGPT.


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
ggplot(data=GSS, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot()+
 xlab("DEGREE") +
 ylab("WORDSUM") +
 ggtitle("Graphic Descriptive Results")
```


Those with Graduate have highest MOSTSUM scoring at 12. Those with a Bachelor's WORDSUM score 10. Those with Highn School education score about 9. Those with Junior college education score 10. Also those with LT High School score about 8.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSS)
```

Those with a Bachelor's degree on average score 10 WORDSUM. Those with Graduate degree score on average 11 WORDSUM. Those with High School education score on average 9. Those who have a Junior College education averagely score 10. Those who have a LT High School score 8 WORDSUM on average. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSS)
summary.aov(model1)
```

Those with a degree on average score WORDSUM about 24, while on the other hand Residuals score approximately 4. P value is less than 0.005 therefore we reject null hypothesis.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The question was is there a difference between the average score on the WORDSUM test and level of schooling DEGREE. P value is less than 0.05 therefore we reject the null hypothesis.

